All Graphs Lead to Rome: Learning Geometric and Cycle Consistent Representations with Graph CNNs

### 2018-11-14 cvpr2019/main.tex
Best viewed in color
Added new figure and a few more cleanups

% Photometric losses make the brightness constancy assumption, which is not as robust as feature matching.
\begin{figure}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyBasic.pdf}
\end{center}
  \caption{
    Example of multi-image matching, with images from Rome16K dataset \cite{li2010location}.
    In this example, an error in matching leads to an inconsistent cycle, ideally these pairwise matches would be consistent.
    In this work, we show how to learn this multi-view matching problem using cycle consistency.
  }
\label{fig:cycconsistex}
\label{fig:onecol}
\end{figure}

@@

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyMainFigure.pdf}
\end{center}
  \caption{
    An outline of the pipeline of this work.
    The graph of matches is put into a Graph Convolutional Neural Network (GCN) \cite{kipf2016semi} and then outputs a low rank embedding of the Adjacency Matrix of the graph.
    The GCN operates on an embedding over the nodes the graph.
    In the figure, the colors on the nodes of the graph in the GCN represent the values of the embedding.
    The final embedding is used to construct a pairwise similarity matrix, which ideally should be a low dimensional cycle consistent representation of the graph adjacency matrix.
    We train the network using a reconstruction loss on the similarity matrix.
    In addition we can use geometric information at to assist training the network, even if we do not have that geometric information at testing time such as epipolar constraints on the point locations or higher order geometric constraints.
  }
\label{fig:pipeline}
\label{fig:onecol}
\end{figure*}

@@

\subsection{Graph Neural Networks}
As input to our method we are given the graph Laplacian $\mat{L} \in \bR^{n \times n}$ to encode the graph structure:
\begin{equation}
  \mat{L} =\; \mat{I} - \mat{D}^{-1/2} \mat{A} \mat{D}^{-1/2} \\
\end{equation}
and an initial embedding $\mat{E}_0 \in \bR^{n \times m_0}$.
Many older methods assume $\mat{L}$ is known a-priori \cite{bruna2013spectral}, and thus can encode graph convolutions using the eigenvectors of $L$.
We do not have this luxury, and thus we use more modern non-spectral methods GNNs:
\begin{equation}
\mat{E}_{i+1} =\; P\left(\sigma\left( \sum_k \mat{L}^k \mat{E}_i \mat{W}_{i,k} \right)\right)
\end{equation}
The maximum degree of $\mat{L}$ determines the number of hops away from a node one layer can access.
Here $\sigma$ is the non-linearity (typically a ReLU), $\mat{W}_{k,i}$ are the learned weights of layer $i$,  and $P$ is the `pooling' operation, known as graph coarsening \cite{bronstein2017geometric, gama2018mimo}.
In the feature matching problem, we require an embedding for each vertex.
Therefore we cannot perform graph coarsening, and thus we only perform the non-linearity step.
We instead use a modification of the approach in Kipf and Welling \cite{kipf2016semi}:
\begin{align}
      \widetilde{\mat{L}} =&\; (\mat{D} + \mat{I})^{-1/2} (\mat{A} + \mat{I}) (\mat{D} + \mat{I})^{-1/2} \\
\mat{E}_{l+1} =&\; \sigma\left(\widetilde{\mat{L}} \mat{E}_i \mat{W}_i \right)  \label{eq:graph_conv}
\end{align}

@@

\subsection{Cycle Consistency}
Let $\mat{M}$ be the noiseless set of matches between our features, with $\mat{M}_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then we can say that, for all $i, j, k$:
\begin{equation}
\mat{M}_{ij} = \mat{M}_{ik} \mat{M}_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stays the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle concistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to.
Then, one can match the $i^{th}$ set of features to the universe using a matrix $\mat{X}_i$.
Then the cycle consistency becomes:
\begin{equation}
\mat{M}_{ij} = \mat{X}_{i}\mat{X}_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.8\linewidth]{figures/GeometricConsistency.pdf}
\end{center}
  \caption{
    Illustrated here is an example of how the geometric loss on one feature.  
    We compute the errors on each feature match to determine their validity.
    Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
  }
\label{fig:cycconsistex}
\label{fig:onecol}
\end{figure}

\subsection{Geometric Side Information}
One of the main advantages of this approach over more traditional optimization based approaches is the ability to add geometric side information into the loss at training time, even if it is not available at test time.
The simplest way to add geometric side losses, and the approach we use here, is to use the epipolar constraint.
The epipolar constraint describes how the positions of features in different images corresponding to the same point should be related.
Given a relative pose $(R_{ij}, T_{ij})$ between two cameras $i$ and $j$  (transforms $j$ to $i$) the epipolar on corresponding feature locations $X_i$ and $X_j$:
\begin{equation}
X_{i}^\top \cross{T_{ij}}R_{ij} X_{j} = 0
\label{eq:essential_constraint_rel}
\end{equation}
In this work we use the two pose epipolar constraint \cite{tron2014quotient}:
\begin{equation}
X_{i}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j} = 0
\label{eq:essential_constraint}
\end{equation}
% Where the $(R_k, T_k)$ are the poses of cameras $i$ and $j$ respectively.
The constraint assumes that the $X_k$ are calibrated i.e. the camera intrinsics are known. 
Given the matrix of ground truth correspondences $\mat{M}_{ij}$ between camera $i$ and $j$, we can formulate this as a loss:
\begin{equation}
\mathcal{L}_{ij,geom}(\mat{M}_{ij}) = \sum_{k,l} (\mat{M}_{ij})_{kl} \left|X_{i,k}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j,l}\right|
\label{eq:geom_cost}
\end{equation}
For our purposes, since we use low rank embeddings $\mat{E}_{i}$, $\mat{E}_{j}$, the loss would look like:
\begin{align}
\mathcal{L}_{geom}(\mat{E})
=&\; \mathrm{tr}(\mat{G}^\top \mat{E}\mat{E}^\top) = \sum_{i,j} (\mat{E})_{i} \cdot (\mat{E})_{j} (\mat{G})_{ij} \\
(\mat{G})_{ij} =&\; \left|X_{i}^\top R_{c(i)}^\top \cross{T_{c(j)} - T_{c(i)}}R_{c(j)} X_{j}\right| \nonumber
\label{eq:geom_cost2}
\end{align}
Where $c(i)$ is the appropriate camera for point index $i$.
In future work we will invistigate using higher order constraints, such as the trifocal tensor, as additional loss terms.

>>>>

% Photometric losses make the brightness constancy assumption, which is not as robust as feature matching.
\begin{figure}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyBasic.pdf}
\end{center}
  \caption{
    Example of multi-image matching, with images from Rome16K dataset \cite{li2010location}.
    In this example, an error in matching leads to an inconsistent cycle, ideally these pairwise matches would be consistent.
    In this work, we show how to learn this multi-view matching problem using cycle consistency.
    Best viewed in color.
  }
\label{fig:cycconsistex}
\label{fig:onecol}
\end{figure}

@@

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyMainFigure.pdf}
\end{center}
  \caption{
    An outline of the pipeline of this work.
    The graph of matches is put into a Graph Convolutional Neural Network (GCN) \cite{kipf2016semi} and then outputs a low rank embedding of the Adjacency Matrix of the graph.
    The GCN operates on an embedding over the nodes the graph.
    In the figure, the colors on the nodes of the graph in the GCN represent the values of the embedding.
    The final embedding is used to construct a pairwise similarity matrix, which ideally should be a low dimensional cycle consistent representation of the graph adjacency matrix.
    We train the network using a reconstruction loss on the similarity matrix.
    In addition we can use geometric information at to assist training the network, even if we do not have that geometric information at testing time such as epipolar constraints on the point locations or higher order geometric constraints.
    Best viewed in color.
  }
\label{fig:pipeline}
\label{fig:onecol}
\end{figure*}

@@

\subsection{Graph Neural Networks}
As input to our method we are given the graph Laplacian $\mat{L} \in \bR^{n \times n}$ to encode the graph structure:
\begin{equation}
  \mat{L} =\; \mat{I} - \mat{D}^{-1/2} \mat{A} \mat{D}^{-1/2} \\
\end{equation}
and an initial embedding $\mat{E}_0 \in \bR^{n \times m_0}$.
Many older methods assume $\mat{L}$ is known a-priori \cite{bruna2013spectral}, and thus can encode graph convolutions using the eigenvectors of $L$.
We do not have this luxury, and thus we use non-spectral GNNs:
\begin{equation}
\mat{E}_{i+1} =\; P\left(\sigma\left( \sum_k \mat{L}^k \mat{E}_i \mat{W}_{i,k} \right)\right)
\end{equation}
The maximum degree of $\mat{L}$ determines the number of hops away from a node one layer can access.
Here $\sigma$ is the non-linearity (typically a ReLU), $\mat{W}_{k,i}$ are the learned weights of layer $i$,  and $P$ is the `pooling' operation, known as graph coarsening \cite{bronstein2017geometric, gama2018mimo}.
In the feature matching problem, we require an embedding for each vertex.
Therefore we cannot perform graph coarsening, and thus we only perform the non-linearity step.
We instead use a modification of the approach in Kipf and Welling \cite{kipf2016semi}:
\begin{align}
      \widetilde{\mat{L}} =&\; (\mat{D} + \mat{I})^{-1/2} (\mat{A} + \mat{I}) (\mat{D} + \mat{I})^{-1/2} \\
\mat{E}_{l+1} =&\; \sigma\left(\widetilde{\mat{L}} \mat{E}_i \mat{W}_i \right)  \label{eq:graph_conv}
\end{align}

@@

\subsection{Cycle Consistency}

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures/UniverseOfFeatures.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    Therefore, we can get the points to have a cycle consistent embedding by mapping each feature in each image to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are much fewer 3D points and thus this corrsponds to a low rank factorization of the corrspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}

Let $\mat{M}$ be the noiseless set of matches between our features, with $\mat{M}_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then we can say that, for all $i, j, k$:
\begin{equation}
\mat{M}_{ij} = \mat{M}_{ik} \mat{M}_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stays the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle concistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a matrix $\mat{X}_i$.
Then the cycle consistency becomes:
\begin{equation}
\mat{M}_{ij} = \mat{X}_{i}\mat{X}_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

\subsection{Geometric Side Information}

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.8\linewidth]{figures/GeometricConsistency.pdf}
\end{center}
  \caption{
    Illustrated here is an example of how the geometric loss on one feature.  
    We compute the errors on each feature match to determine their validity.
    Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
    Best viewed in color.
  }
\label{fig:cycconsistex}
\label{fig:onecol}
\end{figure}

One of the main advantages of this approach over more traditional optimization based approaches is the ability to add geometric side information into the loss at training time, even if it is not available at test time.
The simplest way to add geometric side losses, and the approach we use here, is to use the epipolar constraint.
The epipolar constraint describes how the positions of features in different images corresponding to the same point should be related.
Given a relative pose $(R_{ij}, T_{ij})$ between two cameras $i$ and $j$  (transforms $j$ to $i$) the epipolar on corresponding feature locations $X_i$ and $X_j$:
\begin{equation}
X_{i}^\top \cross{T_{ij}}R_{ij} X_{j} = 0
\label{eq:essential_constraint_rel}
\end{equation}
In this work we use the two pose epipolar constraint \cite{tron2014quotient}:
\begin{equation}
X_{i}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j} = 0
\label{eq:essential_constraint}
\end{equation}
% Where the $(R_k, T_k)$ are the poses of cameras $i$ and $j$ respectively.
The constraint assumes that the $X_k$ are calibrated i.e. the camera intrinsics are known. 
Given the matrix of ground truth correspondences $\mat{M}_{ij}$ between camera $i$ and $j$, we can formulate this as a loss:
\begin{equation}
\mathcal{L}_{ij,geom}(\mat{M}_{ij}) = \sum_{k,l} (\mat{M}_{ij})_{kl} \left|X_{i,k}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j,l}\right|
\label{eq:geom_cost}
\end{equation}
For our purposes, since we use low rank embeddings $\mat{E}_{i}$, $\mat{E}_{j}$, the loss would look like:
\begin{align}
\mathcal{L}_{geom}(\mat{E})
=&\; \mathrm{tr}(\mat{G}^\top \mat{E}\mat{E}^\top) = \sum_{i,j} (\mat{E})_{i} \cdot (\mat{E})_{j} (\mat{G})_{ij} \\
(\mat{G})_{ij} =&\; \left|X_{i}^\top R_{c(i)}^\top \cross{T_{c(j)} - T_{c(i)}}R_{c(j)} X_{j}\right| \nonumber
\label{eq:geom_cost2}
\end{align}
Where $c(i)$ is the appropriate camera for point index $i$.
In future work we will invistigate using higher order constraints, such as the trifocal tensor, as additional loss terms.
